### 2020-01-28 paper/shared.tex
Fix arxiv refs and subject codes and figures

\usepackage{amssymb, amsmath}
\usepackage{algorithmic}
\usepackage{cleveref}
\usepackage{graphicx}
\usepackage[caption=false]{subfig}

% Prevent itemized lists from running into the left margin inside theorems and proofs
\usepackage{enumitem}
\setlist[enumerate]{leftmargin=.5in}
\setlist[itemize]{leftmargin=.5in}

% Used for creating new theorem and remark environments
\newsiamremark{note}{Note}
\crefname{note}{Note}{Notes}

% Sets running headers as well as PDF title and authors
\headers{Multifidelity ABC-SMC}{T.P. Prescott and R.E. Baker}

% Title. If the supplement option is on, then "Supplementary Material"
% is automatically inserted before the title.
\title{Multifidelity Approximate Bayesian Computation with Sequential Monte Carlo Parameter Sampling\thanks{
Submitted to the editors January 2020 \funding{
This work is supported by BBSRC through grant BB/R000816/1.}}}

\author{
Thomas P. Prescott\thanks{Mathematical Institute, University of Oxford, Woodstock Road, Oxford, OX2 6GG, UK (\email{prescott@maths.ox.ac.uk, ruth.baker@maths.ox.ac.uk})\funding{REB is supported by a Royal Society Wolfson Research Merit
Award and a Leverhulme Research Fellowship.}}
\and Ruth E. Baker\footnotemark[2]
}

\newcommand{\obs}[1]{#1_{\mathrm{obs}}}
\newcommand{\tp}[1]{#1_{\mathrm{tp}}}
\newcommand{\fp}[1]{#1_{\mathrm{fp}}}
\newcommand{\fn}[1]{#1_{\mathrm{fn}}}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}

\graphicspath{{figs/out/png/}}